Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2021-05-15 paper/case-study.tex
Full pass done. Spell check. Minor edits. Paper fits.

In this section, we apply the BriefCASE tool to the development of a UAV surveillance system in which a UAV receives commands from a ground station to conduct surveillance over a geographical region. The on-board mission computer then generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission. The UAV is also given a set of \textit{keep-in} and \textit{keep-out} zones that may constrain its flight path.

@@

Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.  We perform a formal analysis using the AGREE tool (which is integrated with the BriefCASE environment) to verify that the model satisfies the contracts.  For the initial version of our design, the verification passes.
%
Although we are satisfied with the results of the formal verification using AGREE, we have not yet analyzed the design for cyber-vulnerabilities.  
In BriefCASE, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.  The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the design.  To satisfy these requirements we need to mitigate the vulnerabilities discovered by the analysis by modifying the design.
%
For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring unverified or malicious code (that could potentially be embedded in the component) will not impact other processes.

In total, seven cyber requirements are generated and imported into our model.  These include four \textit{well-formedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.  Requirements are imported into the model as goals in a Resolute assurance case.  Because we can run Resolute at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.

The intent of the \textit{well-formedness} requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.  In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.  By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.  The Filter transform is therefore applied for each well-formedness requirement, inserting filter components on the incoming and outgoing UxAS connections.

The filter behavior for each component is specified in AGREE.  Not only does this enable formal verification within the modeling environment, but it also provides a means for synthesizing the component implementation in a provably correct manner using the SPLAT tool.  Because SPLAT is integrated with BriefCASE, the proof it emits when synthesizing component code is used as evidence in the Resolute goal for the corresponding mitigation.  When Resolute evaluates whether such a goal is supported by evidence, it checks for the existence of the synthesis proof in addition to verifying that the architecture is correct.

@@

In addition to monitoring the UxAS output for malformed messages, we must also monitor for suspicious behavior.  This requires adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.  The Monitor transform is applied for this class of mitigation.  In general, monitors observe a channel and compare its contents against a reference signal or constant.  The monitor policy specifies acceptable comparisons, and if violated, the monitor sends out an alert.  A monitor can choose to \textit{gate} the observed signal, in which case it also acts as a special kind of filter and drops the message if the policy is violated.  The \textit{monitoring} requirements drive two transforms.  The first adds a \textit{response} monitor to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.  The second adds a \textit{geofence} monitor to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.  The Geofence Monitor is a gated monitor; it prevents the observed Automation Response message from reaching the Waypoint Manager.

Similar to the filters, the monitor policies are specified in AGREE. For example, 
%The Response Monitor specification is shown in \figref{fig:response-monitor}, and 
the Geofence Monitor specification is shown in \figref{fig:geofence-monitor}.

@@

The requirements that have been addressed so far mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.  But we also want to protect against a compromised Ground Station that could potentially transmit well-formed, but malicious commands. The final cyber-requirement is mitigated by the Attestation transform~\cite{attestation-copland}, which adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.  The Attestation Manager is implemented in CakeML and automatically inserted into the application code base by BriefCASE.  Because the Attestation Gate acts as a filter, the transform automatically generates its complete AGREE specification.%, as shown in \figref{fig:attestation}.

@@

The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicate with each other over Ethernet.  The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.  The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.  Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior. In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation transform.

We first performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.  
A third party evaluator repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful.
The following scenarios were exercised (status messages from the high-assurance components in the hardened system are shown in \figref{fig:mitigation-output}):

\paragraph{Infected Ground Station} In this scenario, one of the application files is modified on the Ground Station, which sends the UAV on a trajectory that violates mission requirements. 
On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.

\paragraph{Malformed Line Search Task message} In this scenario, the Line Search Task message contains a waypoint with a longitude value outside the permitted range.  Such malformed messages could exploit vulnerabilities in the on-board software.  
This vulnerability is mitigated by inserting a well-formedness filter.  The filter prevented the Line Search Task message from reaching UxAS.

\paragraph{UxAS vulnerability exploit} In this scenario, Line Search Tasks with greater than 90 waypoints were transmitted from the Ground Station, which triggered a vulnerability that crashes UxAS.   When this occurs, UxAS is prevented from generating an Automation Response.  
This vulnerability is mitigated by inserting a Response Monitor, which checks to see that UxAS outputs an Automation Response message shortly after receiving an Automation Request.  For this scenario, we chose for the monitor to output a status message, which would then be received by the Ground Station and an appropriate action taken.

\paragraph{UxAS trojan modifies flight plan} In this scenario, a trojan embedded in UxAS attempts to cause the UAV to fly into the specified keep-out zone by modifying the mission command waypoints in the Automation Response.  
On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and instead returned the UAV to Home Base.

>>>>

In this section, we outline the application of the BriefCASE tool towards the development of a UAV surveillance system as a part of the DARPA CASE program.
%apply the BriefCASE tool towards the development of a UAV surveillance system in which 
The UAV receives commands from a ground station to conduct surveillance over a geographical region. In response, the UAV's on-board mission computer generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission. The UAV is also given a set of \textit{keep-in} and \textit{keep-out} zones that may constrain its flight path.

@@

Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.  We perform a formal analysis using the AGREE tool (which is integrated with the BriefCASE environment) to verify that the model satisfies the contracts.  For the initial version of our design, the verification passes.
%
Although we are satisfied with the results of the formal verification using AGREE, we have not yet analyzed the design for cyber-vulnerabilities.  
In BriefCASE, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.  The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the design,
and we need to satisfy these requirements by modifying the design.
%  To satisfy these requirements we need to mitigate the vulnerabilities discovered by the analysis by modifying the design.
%
For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) will not impact other processes.

In total, seven cyber requirements are generated and imported into our model.  These include four \textit{wellformedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.  Requirements are imported into the model as goals in a Resolute assurance case.  Because we can run Resolute at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.

The intent of the \textit{wellformedness} requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.  In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.  By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.  The Filter transform is therefore applied for each wellformedness requirement, inserting filter components on the incoming and outgoing UxAS connections.

The filter behavior for each component is specified in AGREE.  This enables formal verification within the modeling environment, and also provides a means for synthesizing the component implementation in a provably correct manner using the SPLAT tool.  Because SPLAT is integrated with BriefCASE, the proof it emits when synthesizing component code is used as evidence in the Resolute goal for the corresponding mitigation.  When Resolute evaluates whether such a goal is supported by evidence, it checks for the existence of the synthesis proof in addition to verifying that the architecture is correct.

@@

In addition to monitoring the UxAS output for malformed messages, we must also monitor for suspicious behavior.  This requires adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.  The Monitor transform is applied for this class of mitigation.  In general, monitors observe a channel and compare its contents against a reference signal or constant.  The monitor policy specifies acceptable comparisons, and if violated, the monitor sends out an alert.  A monitor can choose to \textit{gate} the observed signal, in which case it also acts as a special kind of filter and drops the message if the policy is violated.  The \textit{monitoring} requirements drive two transforms.  The first adds a \textit{response} monitor to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.  The second adds a \textit{geofence} monitor to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.  The Geofence Monitor is a gated monitor; it prevents the observed Automation Response message from reaching the Waypoint Manager. 
%
Similar to the filters, the monitor policies are specified in AGREE. For example, 
%The Response Monitor specification is shown in \figref{fig:response-monitor}, and 
the Geofence Monitor specification is shown in \figref{fig:geofence-monitor}.

@@

So far, we have addressed requirements that 
%The requirements that have been addressed so far
mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.  But we also want to protect against a compromised Ground Station transmitting wellformed, but malicious commands. The final cyber requirement is mitigated by the Attestation transform~\cite{attestation-copland} that adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.  The Attestation Manager is implemented in CakeML and automatically inserted into the application code base by BriefCASE.  Because the Attestation Gate acts as a filter, the transform automatically generates its complete AGREE specification.%, as shown in \figref{fig:attestation}.

@@

The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicated with each other over ethernet.  The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.  The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.  Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior. In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation transform.
%
We performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.  
A third party evaluator repeated the tests on the hardened system and % were able to 
demonstrated that our mitigations were successful.
The following scenarios were exercised (status messages from the high-assurance components in the hardened system are in \figref{fig:mitigation-output}):

\paragraph{Infected Ground Station} An application file is modified on the Ground Station, which sends the UAV on a mission-violating trajectory.
On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.

\paragraph{Malformed Line Search Task message} The Line Search Task message contained a waypoint with a longitude outside the permitted range to exploit vulnerabilities in the on-board software.  The wellformedness filter for Line Search Task messages prevented this message from reaching the UxAS.

\paragraph{UxAS vulnerability exploit} Line Search Tasks with greater than 90 waypoints were transmitted from the Ground Station, triggering a vulnerability that crashes UxAS.   When this occurs, UxAS is unable to generate an Automation Response.  
This vulnerability is mitigated by inserting a Response Monitor that checks to see if UxAS outputs an Automation Response message shortly after receiving an Automation Request.  For this scenario, we chose for the monitor to output a status message, which would then be received by the Ground Station and an appropriate action taken.

\paragraph{UxAS trojan modifies flight plan} A trojan embedded in UxAS attempts to cause the UAV to fly into a specified keep-out zone by modifying the mission command waypoints in the Automation Response.  
On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and returned the UAV to Home Base.

### 2021-05-15 paper/conclusion.tex
Full pass done. Spell check. Minor edits. Paper fits.

The DARPA CASE program is creating tools for systems engineers to integrate cyber-vulnerability analysis and mitigation into their development workflow. The resulting BriefCASE tool suite includes analysis tools for generating cyber requirements, cyber resiliency tools for addressing the requirements, verification tools for ensuring design correctness, and synthesis tools for generating provably correct code. Several of the BriefCASE transforms (filter, monitor, gate) insert components into the model whose behavior can be formally specified using the AGREE language.
The SPLAT tool can then automatically generate CakeML implementations for these components, along with proofs of correctness for assurance that the implementation satisfies the specification.

BriefCASE was applied to a full-scale case study using the Air Force Research Laboratory's OpenUxAS software, exercising a range of built-in cyber resiliency mitigations to meet cyber-requirements. 
The size and scale of the study suggests BriefCASE meets the complexity demands of real-world design.

We are currently in the process of applying BriefCASE to the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the DARPA CASE program.  Other ongoing work includes adding support for uninterpreted functions, mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.

>>>>

The DARPA CASE program is creating tools for systems engineers to
integrate cyber-vulnerability analysis and mitigation into their
development workflow. The resulting BriefCASE tool suite includes
analysis tools for generating cyber requirements, cyber resiliency
tools for addressing the requirements, verification tools for ensuring
design correctness, and synthesis tools for generating provably
correct code. Several of the BriefCASE transforms (filter, monitor,
gate) insert components into the model whose behavior can be formally
specified using the AGREE language.  The SPLAT tool can then
automatically generate CakeML implementations for these components,
along with proofs of correctness for assurance that the implementation
satisfies the specification.
%
BriefCASE was applied to a full-scale case study using the Air Force
Research Laboratory's OpenUxAS software, exercising a range of
built-in cyber resiliency mitigations to meet cyber-requirements.  The
size and scale of the study suggests BriefCASE meets the complexity
demands of real-world design.

We are currently in the process of applying BriefCASE to the design of
an application using the Collins Common Avionics Architecture System
(CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the
DARPA CASE program.  Other ongoing and future work includes
adding support for uninterpreted functions, mechanizing the
correctness-of-synthesis proofs in HOL4, and lifting the proof results
to infinite streams.

### 2021-05-15 paper/example.tex
Full pass done. Spell check. Minor edits. Paper fits.

\begin{figure*}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.4]{example.png} \\
      (a) \\
      \includegraphics[scale=0.4]{example-certificate.png} \\
      (b)
    \end{tabular}
  \end{center}
\caption{Automated UAV route planning system. (a) Unhardened system. (b) Failure certificate.}
\label{fig:example}
\end{figure*}

@@

The initial AGREE contract specifications for the components and the overall system make no assumptions about the integrity of the inputs and outputs. A cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously. In response, the developer modifies the AGREE specification for the AI route planner to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output.  In other words, the AI output is unconstrained in the AGREE specification, allowing it to take on any value and allowing the AI component to send that value at any time.

The contract specifications for the other components are also updated with the threat analysis information by adding in new requirements to mitigate the identified cyber-vulnerabilities. For example, a designer adds to the specification for the legacy waypoint manager assumptions about its input being \emph{well-formed} since that is no longer known \emph{a priori} as the AI route planner output is unconstrained. The systems designer is also responsible for defining the meaning of ``well-formed''; in this example, it is a predicate checking that the waypoints are within bounded value ranges for latitude, longitude, and altitude, with an additional integrity check on message IDs.

The developer adds two other requirements to the SW AGREE specification related to the cyber-vulnerability analysis.  The first, \emph{Waypoint is well-formed} requires all the waypoints sent to the UAV flight controller to be well-formed (it detects if malformed waypoints are propagated downstream).  The second, \emph{Alert if start is not bounded relative to a request}, requires an automation request to correspond with an automation response either in the same step or within one step (it detects if a mission is being started without a request or if the start is delayed more than one step after a request). The goal of these two requirements is to detect if the untrusted component is trying to prevent the UAV from flying its intended mission or to fly a wrong mission.

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.3]{hardened.png} \\
      (a) \\
      \includegraphics[scale=0.4]{hardened-certificate.png} \\
      (b)
    \end{tabular}
  \end{center}
  \caption{Hardened UAV system. (a) The implementation with high-assurance components. (b) Passing certificate.}
  \label{fig:hardened}
\end{figure*}

>>>>

\begin{figure*}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.3]{example.png} \\
      (a) \\
      \includegraphics[scale=0.3]{example-certificate.png} \\
      (b)
    \end{tabular}
  \end{center}
\caption{Automated UAV route planning system. (a) Unhardened system. (b) Failure certificate.}
\label{fig:example}
\end{figure*}

@@

The initial AGREE contract specifications for the components and the overall system make no assumptions about the integrity of the inputs and outputs. A cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously. In response, the system designer modifies the AGREE specification for the AI route planner to model this ability to behave maliciously
(as an untrusted component)
by removing any guarantees about its output.  In other words, the AI output is unconstrained in the AGREE specification, allowing it to take on any value and allowing the AI component to send that value at any time.

The contract specifications for the other components are also updated with the threat analysis information by adding in new requirements to mitigate the identified cyber-vulnerabilities. For example, a designer adds to the specification for the legacy waypoint manager assumptions about its input being \emph{well-formed} since that is no longer known \emph{a priori} as the AI route planner output is unconstrained. The systems designer is also responsible for defining the meaning of well-formed; in this example, it is a predicate checking that the waypoints are within bounded value ranges for latitude, longitude, and altitude, with an additional integrity check on message IDs.

The designer adds two other requirements to the SW AGREE specification related to the cyber-vulnerability analysis.  The first, \emph{Waypoint is well-formed} requires all the waypoints sent to the UAV flight controller to be well-formed (it detects if malformed waypoints are propagated downstream).  The second, \emph{Alert if start is not bounded relative to a request}, requires an automation request to correspond with an automation response either in the same step or within one step (it detects if a mission is being started without a request or if the start is delayed more than one step after a request). The goal of these two requirements is to detect if the untrusted component is trying to prevent the UAV from flying its intended mission or to fly a wrong mission.

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.3]{hardened.png} \\
      (a) \\
      \includegraphics[scale=0.3]{hardened-certificate.png} \\
      (b)
    \end{tabular}
  \end{center}
  \caption{Hardened UAV system. (a) The implementation with high-assurance components. (b) Passing certificate.}
  \label{fig:hardened}
\end{figure*}

### 2021-05-15 paper/related-work.tex
Full pass done. Spell check. Minor edits. Paper fits.

Assume-guarantee reasoning for compositional verification in reactive systems is well-studied \cite{10.1007/978-3-642-28891-3_13, agree2013, 10.1145/2658982.2527272, 10.1007/978-3-319-17524-9_7}. Automated proofs of realizability for assume-guarantee reasoning are useful for engineers implementing components in the system \cite{10.1007/978-3-319-17524-9_13, 10.1007/978-3-319-29613-5_7}. Algorithms for actual component synthesis for Lustre models using k-induction or IC3/PDR provide an automated path from the assume-guarantee reasoning to an actual satisfying node implementation \cite{katis2017synthesis, 10.1007/978-3-319-89963-3_10}. These synthesis algorithms generate code in the Lustre modeling language but do not provide a path to a low-level implementation that could be fielded.

Contracts are similar to assume-guarantee reasoning but are targeted to programming languages. Contracts are often more expressive than assume-guarantee reasoning being stateful and higher-order \cite{10.1145/583852.581484}. As contracts are often written in the target language, synthesis is not a problem for monitoring but comes with significant overhead \cite{10.1007/978-3-642-28869-2_11}. A monitor for a contract can be removed when it can be statically proven that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.

>>>>

Assume-guarantee reasoning for compositional verification in reactive systems is well-studied \cite{10.1007/978-3-642-28891-3_13, agree2013, 10.1145/2658982.2527272, 10.1007/978-3-319-17524-9_7}. Automated proofs of realizability for assume-guarantee reasoning are useful for engineers implementing components in the system \cite{10.1007/978-3-319-17524-9_13, 10.1007/978-3-319-29613-5_7}. Algorithms for %actual
component synthesis for Lustre models using k-induction or IC3/PDR provide an automated path from the assume-guarantee reasoning to an actual satisfying node implementation \cite{katis2017synthesis, 10.1007/978-3-319-89963-3_10}. These algorithms synthesize code in the Lustre modeling language but do not provide a path to a low-level implementation that could be fielded.

Contracts are similar to assume-guarantee reasoning except they target programming languages. They are often more expressive than assume-guarantee reasoning being stateful and higher-order \cite{10.1145/583852.581484}. As contracts are often written in the target language, synthesizing monitors is not a problem but comes with %significant 
overhead \cite{10.1007/978-3-642-28869-2_11}. A monitor for a contract can be removed when it can be statically proven that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.

### 2021-05-15 paper/synthesis.tex
Full pass done. Spell check. Minor edits. Paper fits.

A filter is intended to be simple, although it may make deep semantic
checks. A filter has one input port and one output; messages on the
input that the filter policy admits pass unchanged to the output port;
all others are dropped (not passed on). We have investigated two kinds
of filter. In the first, a relatively shallow scan of the input
suffices to enforce the policy. For example, we have used the
expressive power of Contiguity Types \cite{contiguity-types} to
enforce \emph{lightweight} bounds constraints on GPS coordinates in
UxAS messages. On the other hand, a filter may need to parse the input
buffer into a data structure specified in AGREE and apply a
user-defined \emph{well-formedness} property, also specified in AGREE,
to the data. Arbitrarily complex well-formedness checks can be made in
this way. \figref{fig:filter-spec} shows a combination where the
checking specified by {\small\verb+WELL_FORMED_AUTOMATION_RESPONSE+}
depends on an underlying check specified by the contiguity type
checking bounds on waypoints.

@@

Partiality is an important consideration: steps 2 and 3 above can
fail; the data might not be parseable or the wellformedness
computation could be badly written and fail at runtime. In such cases,
the filter should recover and yield control without passing the input
onwards. In these cases, the filter is behaving as it should, but we
must also guard against situations in which a \emph{correctly
  specified} filter fails at runtime. This kind of defect arises when
the filter \emph{ought} to accept a message, but lack of resources
results in the filter failing to do so. For example, the parse of a
message might need more space than has been allocated; another example
could be if the timeslice provided by the scheduler is too short for
the wellformedness computation to finish. Thus resource bounds need to
be included in the correctness argument.

@@

The contiguity type specification and well-formedness predicate for
the filter are shown in \figref{fig:filter-spec} and the synthesized
CakeML code is in \figref{fig:filter-cakeml}. The code is called at
dispatch by the scheduler. The \texttt{API.callFFI} is the link to the
communication fabric to capture input and provide output. The body of
the function restates the filter contract to make the appropriate
assignments in a way that matches the truth value of the predicate in
the filter guarantee.  The auto-generated AGREE specification raises
an alert output when the relation is violated.

@@

Monitors are intended to track and analyze the externally visible
behavior of system components through time. Therefore, they require
more extensive computational ability than filters. In particular, our
basic notion of a monitor is that it embodies a predicate over its
input and output streams, able to access the value of a stream at any
earlier point in time, if necessary. Monitors commonly use state to
keep track of earlier values, unlike filters which, for us, are
typically stateless components. (However, there is nothing in our
approach that forbids stateful filters: they can be realized by
monitors.) A monitor specification is mapped by code generation to a
state transformation function of the following abstract type:
\[
\konst{stepFn} : \mathit{input} \times \mathit{stateVars} \to \mathit{stateVars} \times \mathit{output}
\]

@@

A monitor may need to accumulate a certain minimum number of
observations before being able to make a meaningful assessment of
behavior. Until that threshold is attained, the monitor is essentially
in a kind of \emph{initialization} phase. In order for correct code to
be generated, monitor specifications need to spell out the values of
output ports when in an initialization phase. For example, suppose a
monitor does some kind of differential assessment of inputs at
adjacent timeslices, alerting when (say) the measured location of a
UAV at times $t$ and $t+1$ is such that the distance between the two
locations is unusually large. Such a monitor needs two measurements
before making its first judgement, but at the time of its first
output, only one measurement will have been made. The specification
must then explicitly state the correct value for the first output.

@@

Intutively, for monitor specification $s$, \konst{stepFn} is the
concrete embodiment of $\konst{SynthEval}\;s$, as defined in Section
\ref{agree-semantics}. Its correctness amounts to showing that, given
a sequence of inputs, and an initial state meeting the initialization
constraints, iterating \konst{stepFn} produces a $\pi \in
\Lang{s}$. Morevover, taking the union over all input sequences and
initial states produces $\Lang{s}$ itself.

>>>>

A filter is intended to be simple, although it may make deep semantic
checks. A filter has one input port and one output; messages on the
input that the filter policy admits pass unchanged to the output port;
all others are dropped (not passed on). We have investigated two kinds
of filter. In the first, a relatively shallow scan of the input
suffices to enforce the policy. For example, we have used the
expressive power of Contiguity Types \cite{contiguity-types} to
enforce \emph{lightweight} bounds constraints on GPS coordinates in
UxAS messages. On the other hand, a filter may need to parse the input
buffer into a data structure specified in AGREE and apply a
user-defined \emph{wellformedness} property, also specified in AGREE,
to the data. Arbitrarily complex wellformedness checks can be made in
this way. \figref{fig:filter-spec} shows a combination where the
checking specified by {\small\verb+WELL_FORMED_AUTOMATION_RESPONSE+}
depends on an underlying check specified by the contiguity type
checking bounds on waypoints.

@@

Partiality is an important consideration: steps 2 and 3 above can
fail; the data might not be parseable or the wellformedness
computation could be badly written and fail at runtime. In such cases,
the filter should recover and yield control without passing the input
onwards. In these cases, the filter is behaving as it should, but we
must also guard against situations in which a \emph{correctly
  specified} filter fails at runtime. This kind of defect arises when
the filter \emph{ought} to accept a message, but lack of resources
results in the filter failing to do so. For example, the parse of a
message might need more space than has been allocated; another example
could be if the time slice provided by the scheduler is too short for
the wellformedness computation to finish. Thus resource bounds need to
be included in the correctness argument.

@@

The contiguity type specification and wellformedness predicate for
the filter are shown in \figref{fig:filter-spec} and the synthesized
CakeML code is in \figref{fig:filter-cakeml}. The code is called at
dispatch by the scheduler. The \texttt{API.callFFI} is the link to the
communication fabric to capture input and provide output. The body of
the function restates the filter contract to make the appropriate
assignments in a way that matches the truth value of the predicate in
the filter guarantee.  The auto-generated AGREE specification raises
an alert output when the relation is violated.

@@

Monitors are intended to track and analyze the externally visible
behavior of system components through time. Therefore, they require
more extensive computational ability than filters. In particular, our
basic notion of a monitor is that it embodies a predicate over its
input and output streams, and is able to access the value of a stream at any
earlier point in time, if necessary. Monitors commonly use state to
keep track of earlier values, unlike filters which, for us, are
typically stateless components. (However, there is nothing in our
approach that forbids stateful filters: they can be realized by
monitors.) A monitor specification is mapped by code generation to a
state transformation function of the following abstract type:
\[
\konst{stepFn} : \mathit{input} \times \mathit{stateVars} \to \mathit{stateVars} \times \mathit{output}
\]

@@

A monitor may need to accumulate a certain minimum number of
observations before being able to make a meaningful assessment of
behavior. Until that threshold is attained, the monitor is essentially
in its \emph{initialization} phase. In order for correct code to
be generated, monitor specifications need to spell out the values of
output ports when in their initialization phases. For example, suppose a
monitor does some kind of differential assessment of inputs at
adjacent time slices, alerting when (say) the measured location of a
UAV at times $t$ and $t+1$ is such that the distance between the two
locations is unusually large. Such a monitor needs two measurements
before making its first judgement, but at the time of its first
output, only one measurement will have been made. The specification
must then explicitly state the correct value for the first output.

@@

Intuitively, for monitor specification $s$, \konst{stepFn} is the
concrete embodiment of $\konst{SynthEval}\;s$, as defined in Section
\ref{agree-semantics}. Its correctness amounts to showing that, given
a sequence of inputs, and an initial state meeting the initialization
constraints, iterating \konst{stepFn} produces a $\pi$ s.t. $\pi \in
\Lang{s}$; and taking the union over all input sequences and
initial states produces $\Lang{s}$ itself.
